DOpElib:\\ Differential Equations and Optimization Environment

### 2013-12-10 Examples/OPT/StatPDE/Example1/content.tex
Removed Typos in Documentation; OPT/StatPDE/Example1 documentation has been augmented; Introduction has been augmented

The example shows how to estimate the error in the cost functional for stationary optimization
problems.

>>>>

\added{
\subsubsection{Background information and program description}
In the following, we describe all extensions to the previous problems 
relevant to solving PDE-based optimization with \dope{}. So far, 
we had only to implement the \texttt{ElementEquation} and the corresponding
matrix \texttt{ElementMatrix}. Now, based on the idea of the reduced cost
functional, we have to compute certain additional equations representing the 
adjoint, tangent, and adjoint hessian equations
\texttt{ElementEquation\underline{ }U, ElementEquation\underline{ }UT,
ElementEquation\underline{ }UTT}
for the state equation and in the same terms arising from the functional
itself. Let us shed some light into all equations by giving some background
information and overview first.

In abstract form, we are given the following optimization problem:
\begin{equation*}
J(q,u) \, \rightarrow \, \min , \quad a(q,u)(\psi) = 0 \quad \forall\psi\in V
\end{equation*}
Lagrangian:
\begin{equation*}
{\cal L}(q,u,z):=J(q,u) - a(q,u)(z)
\end{equation*}
Optimality system (KKT system):
\begin{align*}
a_u'(q,u)(\phi , z) &= J'_u(q,u)(\phi) \quad \forall \phi\in V \\
a_q'(q,u)(\chi , z) &= J'_q(q,u)(\chi) \quad \forall \chi\in Q \\
a(q,u)(\psi) &= 0 \qquad \forall\psi\in V
\end{align*}
or equivalently, in terms of the Lagrangian
\begin{align*}
{\cal L}'_u(q,u,z)(\phi) &= 0 \quad\forall \phi\in V \quad\text{(Adjoint Equation)} \\
{\cal L}'_q(q,u,z)(\chi) &= 0 \quad\forall \chi\in V \quad\text{(Gradient Equation)} \\
{\cal L}'_z(q,u,z)(\psi) &= 0 \quad\forall \psi\in V \quad\text{(State Equation)}
\end{align*}
The continuous problem is discretized by a standard Galerkin method using 
finite dimensional subspaces $Q_h \times V_h \subset Q\times V$:
\begin{equation*}
J(q,u) \, \rightarrow \, \min , \quad a(q,u)(\psi) = 0 \quad \forall\psi\in V
\end{equation*}
Discrete saddle-point problems
\begin{align*}
a_u'(q_h,u_h)(\phi_h , z_h) &= J'_u(q_h,u_h)(\phi_h) \quad \forall \phi_h\in V_h \\
a_q'(q_h,u_h)(\chi_h , z_h) &= J'_q(q_h,u_h)(\chi_h) \quad \forall \chi_h\in Q_h \\
a(q_h,u_h)(\psi_h) &= 0 \qquad \forall\psi_h\in V_h
\end{align*}
\subsubsection{Solution process}
In this section, we briefly discuss the solution process for the 
optimization problem. For further details, we refer to the standard literature.
The unconstrained optimal control problem is reformulated as follows. 
We introduce the solution operator $S:Q\rightarrow V$ of the state equation.
Then:
\begin{equation*}
j(q):= J(q,S(q))\, \rightarrow \, \min , \quad a(q,S(q))(\Psi) = 0 \quad 
\forall \Psi \in V.
\end{equation*}
The local existence and sufficient regularity of $S$ is assumed. The 
necessary optimality conditions of first and second order are
\begin{equation*}
j'(q)(\delta q) = 0 , \quad j''(q)(\delta q,\delta q) \geq 0 \quad 
\forall\delta q\in Q.
\end{equation*} 
The derivatives of the reduced functional can be computed using the 
Lagrangian
\begin{equation*}
{\cal L}(q,u,z) =J(q,u) - a(q,u)(z)
\end{equation*}
as already introduced. Let $q\in Q$, and the corresponding state $u=S(q)\in V$ 
be given. To calculate the derivative of the reduced cost functional $j$, 
we introduce the dual variable
$z\in V$ solving the\\[3mm]
%
\textbf{Dual equation}\\
\begin{equation*}
{\cal L}_U'(q,U,Z)(\Psi) = 0 \quad\forall\Psi\in V.
\end{equation*}
Then,
\begin{equation*}
j'(q)(\delta q) = {\cal L}_q' (q,U,Z)(\delta q) \quad \text{for }\delta q\in Q.
\end{equation*}
\\[2mm]
%
To calculate the second derivatives, we need to solve additional equations.\\
%
Let $\delta q\in Q$ be a given direction. Then we search $\delta U\in V$  
solving the\\[2mm]
%
\textbf{Tangent equation}\\
%
\begin{equation*}
{\cal L}_{qZ}'' (q,U,Z)(\delta q,\Phi) + {\cal L}_{UZ}'' (q,U,Z)(\delta
U,\Phi) = 0
\quad\forall \Phi\in V.
\end{equation*}
Further, we have an auxilliary\\[2mm]  
%
\textbf{Dual for Hessian equation}
to find $\delta Z\in V$ solving
\begin{equation*}
{\cal L}_{qU}'' (q,U,Z)(\delta q,\Phi) + {\cal L}_{UU}'' 
(q,U,Z)(\delta U,\Phi) + {\cal L}_{ZU}'' (q,U,Z)(\delta Z,\Phi) =0 
\quad\forall \Phi\in V.
\end{equation*}

Then, for $\delta r\in Q$, we can express the second derivatives of $j$ by
\begin{align*}
j''(q)(\delta q, \delta r) 
&= {\cal L}_{qq}'' (q,U,Z)(\delta q,\delta r) \\
&\;\;\;\;+ {\cal L}_{Uq}'' (q,U,Z)(\delta U,\delta r) \\ 
&\;\;\;\;+ {\cal L}_{Zq}'' (q,U,Z)(\delta Z,\delta r).
\end{align*}
With these terms, we can calculate the Newton direction $\delta q$, at a given 
iterate $q^n$, as solution to the problem
\begin{align*}
j''(q^n )(\delta q,\chi ) =&\, -j'(q^n)(\chi) \quad\forall\chi \in Q.
\end{align*}

Moreover, we would like to work in the Hilbert space $Q$. However, the derivative 
$j'(q) \in H^*$ only. Hence, we need to calculate the Riesz representation 
for the gradient $\nabla j(q) \in H$ using the definition:
\[
 (\nabla j(q),\delta q)_Q = j'(q)(\delta q) \quad \forall\,\delta q\in Q.
\]
In the given example, $Q = L^2(\Omega)$ and hence the scalar product will be the
standard $L^2$-inner product. Simillarly, we can define the Hessian operator 
$H(q) \in mathcal L(Q,Q)$ by defining
\[
 (H(q)\tau q,\delta q)_Q = j''(q)(\tau q,\delta q) \quad \forall\,\delta q,\tau q\in Q.
\]
}
\todo[inline]{Brauchen wir den n\"achsten Block wirklich? Das steht 1 Seite dar\"uber schon, 
die Notation passt aber nicht immer, z.B. $J''(u)$, fehlende Argumente und Funktionen (z.N. Dual Hessian Equation)
. 
Ansonsten müssen wir den Block nochmal durchgehen um die Sachen zu beheben.}

Finally, we recall the representations of the previous work. First, we recall 
in terms of the Lagrangian:

\textbf{Adjoint:} Find $z\in X$ such that
\begin{equation*}
{\cal L}_u' (q,u,z)(\phi) = 0 \quad \forall\phi\in X. 
\end{equation*}
\textbf{Tangent (linearized state):} Find $\delta u\in X$ such that
\begin{equation*}
{\cal L}_{qz}'' (q,u,z)(\delta q, \phi) +
{\cal L}_{uz}'' (q,u,z)(\delta u, \phi) = 0 \quad \forall\phi\in X. 
\end{equation*}

\textbf{Additional adjoint:} Find $\delta z\in X$ such that
\begin{equation*}
{\cal L}_{qu}'' (q,u,z)(\delta q, \phi) +
{\cal L}_{uu}'' (q,u,z)(\delta u, \phi) +
{\cal L}_{zu}'' (q,u,z)(\delta z, \phi) = 0 \quad \forall\phi\in X. 
\end{equation*}

Especially, the equations for stationary optimization problems are
given as follows:

\textbf{Adjoint:} Find $z\in X$ such that
\begin{equation*}
a_u'(q,u)(\phi,z) = J'(u)(\phi) \quad \forall\phi\in X. 
\end{equation*}
\textbf{Tangent:} Find $\delta u\in X$ such that
\begin{equation*}
a_u'(q,u)(\delta u,\phi) = -a_q'(q,u)(\delta q,\phi)  \quad \forall\phi\in X. 
\end{equation*}
\textbf{Additional adjoint:} Find $\delta z\in X$ such that
\begin{equation*}
a_u'(q,u)(\phi,\delta z) = a_{uu}''(q,u)(\delta u,\phi,z)
-a_{qu}''(q,u)(\delta q,\phi,z) + J''(u)(\delta,\phi) \quad \forall\phi\in X. 
\end{equation*}

Using the Lagrangian, the previous expressions read

\textbf{Lagrangian}
\begin{equation*}
{\cal L}(q,u,z) =J(q,u) - a(q,u)(z)
\end{equation*}

\textbf{State Equation}
\begin{align*}
a(q,u)(\phi) = 0 
\end{align*}

\textbf{Dual Equation}

\begin{align*}
{\cal L}_{u}' (q,u,z)(\phi ) &= 0 \\
\Leftrightarrow a_u' (q,u)(\phi,z) &= J_u' (q,u)(\phi)
\end{align*}

\textbf{Tangent Equation}

\begin{align*}
{\cal L}_{qz}'' (q,u,z)(\delta q, \phi) 
+ {\cal L}_{uz}'' (q,u,z)(\delta u, \phi) &= 0 \\
\Leftrightarrow a_u' (q,u)(\delta u,\phi) + a_q' (q,u)(\delta q,\phi) &= 0
\end{align*}

\textbf{Dual Hessian Equation}

\begin{align*}
{\cal L}_{qu}'' (q,u,z)(\delta q, \phi) 
&+ {\cal L}_{uu}'' (q,u,z)(\delta u, \phi) \\
&+ {\cal L}_{zu}'' (q,u,z)(\delta z, \phi) = 0 \\
\Leftrightarrow a_u' (q,u)(\phi, \delta z) 
&+ a_{uu}' (q,u)(\delta u,\phi, z) \\
&+ a_{qu}' (q,u)(\delta q,\phi, z)  = J_{uu}''(q,u)(\cdot ,\cdot )
\end{align*}

\textbf{Newton left hand side: first derivative}

\begin{align*}
j'(q)(\delta q) &= {\cal L}_{q}' (q,u,z)(\delta q) \quad\text{(Gradient equation)} \\
\Leftrightarrow j'(q)(\delta q) &= -a_q' (q,u)(\delta q,z) + J_{q}'(q,u)(\delta q)
\end{align*}

\textbf{Newton right hand side: second derivative}

\begin{align*}
j''(q)(\delta q, \delta r) &= {\cal L}_{qq}'' (q,u,z)(\delta q, \delta r) + 
{\cal L}_{uq}'' (q,u,z)(\delta u, \delta r) \\
&\quad +  {\cal L}_{zq}'' (q,u,z)(\delta z, \delta r) \\
\Leftrightarrow  j''(q)(\delta q, \delta r) &= 
J_{qq}''(q,u)(\delta q, \delta r) -a_{qq}'' (q,u)(\delta q,\delta r, z) \\
&\quad - a_{uq}'' (q,u)(\delta u,\delta r, z) -a_{q}' (q,u)(\delta r,\delta z)  
\end{align*}
\todo[inline]{end of delete comment}

\subsubsection{Implementation in \dope{}}
\added{
From the previous details, and the definition of the Lagrangian and its 
derivatives it is clear, that the user has to provide the respective derivatives. 
Since the Lagrangian consists of the PDE and the cost functional it 
is sufficient to provide the respective derivatives, while \texttt{DOpElib} 
will assemble them as required.
Test functions for vector valued terms will be denoted by an index $i$ while 
matrix valued terms are indexed in $i$ and $j$. Test functions in the control 
space $Q$ are denoted as $\phi^q_i$ while those in the state space $V$ are denoted 
as $\phi_i$.

To solve the linear equations the following matrices are needed
\begin{align*}
\text{ElementMatrix} \quad 
&\Leftrightarrow \quad a_{i,j} = a'_u(q,u)(\phi_j,\phi_i), \\
\text{ControlElementMatrix } \quad 
&\Leftrightarrow \quad a_{i,j} = (\phi^q_j, \phi^q_i)_Q.
\end{align*}
The first one is required for all primal and dual PDE solves, while the second one is needed 
to calculate the Riesz representation of the derivatives of $j$. 
If desired, the matrix for the adjoint PDEs can be provided separately as 
\texttt{ElementMatrix\underline{ }T}, but otherwise this will be 
calculated automatically from the primal matrix.

Additional terms are needed to calculate the corresponding 
right hand sides. These are for the PDE
\begin{align*}
\text{ElementEquation (state)} \quad &\Leftrightarrow \quad a(q,u)(\phi_i), \\
\text{ElementEquation\underline{ }U (adjoint)} \quad 
&\Leftrightarrow \quad a_u'(q,u)(\phi_i, z), \\ 
\text{ElementEquation\underline{ }UT (tangent)} \quad 
&\Leftrightarrow \quad a_u'(q,u)(\delta u, \phi_i), \\
\text{ElementEquation\underline{ }UTT (adjoint hessian)} \quad 
&\Leftrightarrow \quad  a_u'(q,u)(\phi_i, \delta z), \\
\text{ElementEquation\underline{ }Q (gradient)} \quad 
&\Leftrightarrow \quad a_q'(q,u)(\phi^q_i,z), \\
\text{ElementEquation\underline{ }QT (tangent)} \quad 
&\Leftrightarrow \quad a_q'(q,u)(\delta q, \phi_i), \\
\text{ElementEquation\underline{ }QTT (hessian)} \quad 
&\Leftrightarrow \quad a_{q}'(q,u)(\phi^q_i,\delta z), \\
\text{ElementEquation\underline{ }UU (adjoint hessian)} \quad 
&\Leftrightarrow \quad a_{uu}''(q,u)(\delta u, \phi_i, z), \\
\text{ElementEquation\underline{ }QU (adjoint hessian)} \quad 
&\Leftrightarrow \quad a_{qu}''(q,u)(\delta q, \phi_i, z), \\
\text{ElementEquation\underline{ }UQ (hessian)} \quad 
&\Leftrightarrow \quad a_{uq}''(q,u)(\delta u, \phi^q_i, z),\\
\text{ElementEquation\underline{ }QQ (hessian)} \quad 
&\Leftrightarrow \quad a_{qq}''(q,u)(\delta q, \phi^q_i, z),\\
\text{ElementRightHandSide (state)} \quad 
&\Leftrightarrow \quad f(\phi_i), \\
\text{ControlElementEquation (gradient or hessian)} \quad 
&\Leftrightarrow \quad (\nabla j(q),\phi^q_i)_Q.
\end{align*}
As for PDE problems, it is up to the user to decide if the ElementRightHandSide
is used, or if the terms are included in the ElementEquation.

For the cost functional, we have to provide
\begin{align*}
\text{ElementValue (all)} \quad  & \Leftrightarrow \quad J(q,u), \\
\text{ElementValue\underline{ }U (all)} \quad  & \Leftrightarrow \quad J_u'(q,u)(\phi_i), \\
\text{ElementValue\underline{ }Q (all)} \quad  & \Leftrightarrow \quad J_q'(q,u)(\phi_i), \\
\text{ElementValue\underline{ }UU (all)} \quad & \Leftrightarrow \quad J_{uu}''(q,u)(\delta u, \phi_i), \\
\text{ElementValue\underline{ }QU (all)} \quad & \Leftrightarrow \quad J_{qu}''(q,u)(\delta q, \phi_i), \\
\text{ElementValue\underline{ }UQ (all)} \quad & \Leftrightarrow \quad J_{uq}''(q,u)(\delta u, \phi^q_i), \\
\text{ElementValue\underline{ }QQ (all)} \quad & \Leftrightarrow \quad J_{qq}''(q,u)(\delta q, \phi^q_i).
\end{align*}

Clearly, if the PDE or cost functional contains other terms, such as boundary or face integrals 
corresponding derivatives must be provided as well.

\subsubsection{Back to the specific equations in this example}
We have
\todo[inline]{Wollen wir die alle? Die Terme $a'_u$ sind ja alle Gleich bis auf das 
jeweilige argument...}
\begin{align*}
a(q,u)(\phi) &= (\nabla u, \nabla\phi) - (q + f, \phi),\\
a_u'(q,u)(\phi, z) &= (\nabla\phi, \nabla z),\\
a_u'(q,u)(\delta u, \phi) &= (\nabla\delta u, \nabla\phi),\\
a_u'(q,u)(\phi, \delta z) &= (\nabla\phi, \nabla\delta z),\\
a_q'(q,u)(\delta q, \phi) &= -(z,\psi^q),\\
a_q'(q,u)(\delta q, \phi) &= (\delta q,\phi),\\
a_{q}'(q,u)(\delta q,\delta z) &= - (\delta z,\psi^q).
\end{align*}
%
For the cost functional, we have the following terms:
\begin{align*}
J(q,u) &= \frac{1}{2} \|u-u^d\|^2 + \frac{\alpha}{2}\|q\|^2,\\
J_u'(q,u)(\phi) = (u-u^d, \phi),\\
J_q'(q,u)(\phi) = \alpha (q,\psi^q),\\
J_{uu}''(q,u)(\psi, \phi) = (\delta u, \phi).
\end{align*}
All other terms, specifically mixed terms with $QU$ etc. are zero in this
example.

\subsubsection{\texttt{main.cc}}
Finally, the main file of the optimization examples does not look very much 
different than for pure PDE computations - which is one of the crucial aims 
of our library. 
Here, instead of using a \texttt{pdeproblemcontainer}, we use 
now an \texttt{optproblemcontainer} which can assemble all additionally needed informations,
such as adjoint and tangent PDEs. 
Furthermore, we define \texttt{ReducedNewtonAlgorithm} and 
\texttt{ReducedTrustregion\underline{ }NewtonAlgorithm}
to solve the optimization problem with a linesearch and a trust-region Newton algorithm.
Of course one would be sufficient, but we wanted to show how to change optimization 
solvers easily using \texttt{DOpElib}.

Next, in the body of the main file, we introduce a second FE function for the 
control variable. Then, we define 
a \texttt{COSTFUNCTIONAL}. Finally, the problem is either solved 
by calling \texttt{Alg.Solve(q)} and/or the user might check if the 
derivatives are implemented correctly by calling 
\texttt{Alg.CheckGrads} or \texttt{Alg.CheckHessian}. The latter 
two functionalities are highly recommended to check your
implementation before wondering about your results.

}

### 2013-12-10 Examples/OPT/StatPDE/Example2/content.tex
Removed Typos in Documentation; OPT/StatPDE/Example1 documentation has been augmented; Introduction has been augmented

\subsubsection{General problem description}
This example solves the following pointwise minimization problem
\begin{align*}
\min_{(q,u)\in \mathbb R^3 \times H_0^1(\Omega; \mathbb R^2)} J(q,u) &= \frac{1}{2} \sum_{i=0}^2\abs{(u-\overline u)(x_i)}^2 + \frac{\alpha}{2}\|q\|^2&&\\
\text{s.t. } (\nabla u,\nabla \phi) &= (f(q),\phi) \quad\forall\,\phi \in H^1_0(\Omega; \mathbb R^2)
\end{align*}
on the domain $\Omega = [0,1]^2$, with zero dirichlet boundary conditions and
\begin{itemize}
\item the observation points
\begin{align*}
x_0 = (0.5, 0.5), \quad x_1 = (0.5, 0.25),\quad x_2 = (0.25, 0.25),
\end{align*}
\item the regularization parameter $\alpha = 0$, 
\item the right hand side
\begin{align*}
 f(q) &= q_0 \left(\begin{matrix}2\pi^2  \sin( \pi x) \sin(\pi y)\\0 \end{matrix}\right)\\
      &+ q_1 \left(\begin{matrix}5\pi^2  \sin( \pi x) \sin(2\pi y)\\0 \end{matrix}\right)\\
      &+ q_2 \left(\begin{matrix}0 \\8\pi^2  \sin(2\pi x) \sin(2\pi y)\end{matrix}\right)\\
\end{align*}
\item and the exact solution given by 
\begin{align*}
 \overline{q} &= \bigl(1;0.5;1\bigr)\\
 \overline{u}& = \left(\begin{matrix} \sin( \pi x)( \sin(\pi y)+0.5\sin(2\pi y))\\\sin(2\pi x) \sin(2\pi y) \end{matrix}\right).
\end{align*}
\end{itemize}
\subsubsection{Program description}
\twick{
In contrast to the first example, the control is now a discrete quantity (for
the three observation points) where we use the \texttt{FE\underline{}Nothing}
element to assigne the three controls. Here, the number of components equals
the number of controls. 
In addition, }notice that the cost functional is of mixed type (from our computational point of view), i.e. the first part is a pointfunctional whereas the regularization part requires the evaluation of a domain integral. To handle this, we need a special integrator as well as a special newtonsolver. Additionally, our \texttt{LocalFunctional} returns as his type:
\begin{verbatim}
      string
      GetType() const
      {
        return "point domain";
      }
\end{verbatim}
This brings along that we have not only to implement 
\begin{itemize} 
\item \texttt{ElementValue},
\item \texttt{ElementValue\_U},
\item \texttt{ElementValue\_Q}, 
\item \texttt{ElementValue\_UU},
\item \texttt{ElementValue\_UQ},
\item \texttt{ElementValue\_QU},
\item \texttt{ElementValue\_QQ}
\end{itemize}
 from \texttt{FunctionalInterface}, but also all the aforementioned methods with a preceeding \text{Point} (\texttt{PointValue} etc.).
%FE_Nothing, handle of the control erlaeutern.

>>>>

\subsubsection{General problem description}
This example solves the following pointwise minimization problem
\begin{align*}
\min_{(q,u)\in \mathbb R^3 \times H_0^1(\Omega; \mathbb R^2)} J(q,u) &= \frac{1}{2} \sum_{i=0}^2\abs{(u-\overline u)(x_i)}^2 + \frac{\alpha}{2}\|q\|^2&&\\
\text{s.t. } (\nabla u,\nabla \phi) &= (f(q),\phi) \quad\forall\,\phi \in H^1_0(\Omega; \mathbb R^2)
\end{align*}
on the domain $\Omega = [0,1]^2$, with zero Dirichlet boundary conditions and
\begin{itemize}
\item the observation points
\begin{align*}
x_0 = (0.5, 0.5), \quad x_1 = (0.5, 0.25),\quad x_2 = (0.25, 0.25),
\end{align*}
\item the regularization parameter $\alpha = 0$, 
\item the right hand side
\begin{align*}
 f(q) &= q_0 \left(\begin{matrix}2\pi^2  \sin( \pi x) \sin(\pi y)\\0 \end{matrix}\right)\\
      &+ q_1 \left(\begin{matrix}5\pi^2  \sin( \pi x) \sin(2\pi y)\\0 \end{matrix}\right)\\
      &+ q_2 \left(\begin{matrix}0 \\8\pi^2  \sin(2\pi x) \sin(2\pi y)\end{matrix}\right)\\
\end{align*}
\item and the exact solution given by 
\begin{align*}
 \overline{q} &= \bigl(1;0.5;1\bigr)\\
 \overline{u}& = \left(\begin{matrix} \sin( \pi x)( \sin(\pi y)+0.5\sin(2\pi y))\\\sin(2\pi x) \sin(2\pi y) \end{matrix}\right).
\end{align*}
\end{itemize}
\subsubsection{Program description}
\twick{
In contrast to the first example, the control is now a discrete quantity (for
the three observation points) where we use the \texttt{FE\underline{}Nothing}
element to assign the three controls. Here, the number of components equals
the number of controls. 
In addition, }notice that the cost functional is of mixed type (from our computational point of view), i.e. the first part is a pointfunctional whereas the regularization part requires the evaluation of a domain integral. To handle this, we need a special integrator as well as a special newtonsolver. Additionally, our \texttt{LocalFunctional} returns as his type:
\begin{verbatim}
      string
      GetType() const
      {
        return "point domain";
      }
\end{verbatim}
This brings along that we have not only to implement 
\begin{itemize} 
\item \texttt{ElementValue},
\item \texttt{ElementValue\_U},
\item \texttt{ElementValue\_Q}, 
\item \texttt{ElementValue\_UU},
\item \texttt{ElementValue\_UQ},
\item \texttt{ElementValue\_QU},
\item \texttt{ElementValue\_QQ}
\end{itemize}
 from \texttt{FunctionalInterface}, but also all the aforementioned methods with a preceding \text{Point} (\texttt{PointValue} etc.).
%FE_Nothing, handle of the control erlaeutern.

### 2013-12-10 Examples/OPT/StatPDE/Example3/content.tex
Removed Typos in Documentation; OPT/StatPDE/Example1 documentation has been augmented; Introduction has been augmented

\subsubsection{General problem description}
In this example, we solve a optimization problems from 
fluid dynamcis. The configuration is similar to 
the fluid optimization problem proposed by Roland Becker 
``Mesh adaption for stationary flow control'' (2000).

@@

The rest of the progam is similar to the previous optimization problems where
we formulate the state equation in a weak form $a(v,p)(\phi)$ such that the 
final problem reads
\[
K(q,v,p) \rightarrow \text{min} \quad \text{s.t.} \quad a(q,v,p)(\phi) = 0.
\]

>>>>

\subsubsection{General problem description}
In this example, we solve a optimization problems from 
fluid dynamics. The configuration is similar to 
the fluid optimization problem proposed by Roland Becker 
``Mesh adaption for stationary flow control'' (2000).

@@

The rest of the program is similar to the previous optimization problems where
we formulate the state equation in a weak form $a(v,p)(\phi)$ such that the 
final problem reads
\[
K(q,v,p) \rightarrow \text{min} \quad \text{s.t.} \quad a(q,v,p)(\phi) = 0.
\]

### 2013-12-10 Examples/OPT/StatPDE/Example4/content.tex
Update Doku und Beseitigung kleinerer Doku-typos in zwei main files.
Removed Typos in Documentation; OPT/StatPDE/Example1 documentation has been augmented; Introduction has been augmented

This example solves the minimization problem
\begin{gather*}
\min J(q,u) = \frac{1}{2} \|u-u^d\|^2 + \frac{\alpha}{2}\|q\|^2\\
\text{s.t. } (\nabla u,\nabla \phi) = (f,\phi)\;\;\forall\,\phi \in H^1_0(\Omega; \mathbb R^2)\\
\end{gather*}
on the domain $\Omega = [0,1]^2$. In addition we set the dirichlet data of the state on the boundary as follows
\begin{gather*}
 u_0(0,y) = q_0,\quad u_0(1,y) = q_1,\quad u_0(x,0) = q_2,\quad u_0(x,1) = q_3,\\
 u_1 = q_4^3.
\end{gather*}
The data is chosen as follows:
\begin{align*}
 f &= \left(\begin{matrix}20\pi^2  \sin( \pi x) \sin(\pi y)\\1 \end{matrix}\right)\\
 u^d&= \left(\begin{matrix}\sin( \pi x) \sin(\pi y)*x\\x \end{matrix}\right)
\end{align*}
with $\alpha = 10$.

>>>>

\subsubsection{General problem description}
This example solves the minimization problem
\begin{gather*}
\min J(q,u) = \frac{1}{2} \|u-u^d\|^2 + \frac{\alpha}{2}\|q\|^2\\
\text{s.t. } (\nabla u,\nabla \phi) = (f,\phi)\;\;\forall\,\phi \in H^1_0(\Omega; \mathbb R^2)\\
\end{gather*}
on the domain $\Omega = [0,1]^2$. In addition we set the Dirichlet data of the state on the boundary as follows
\begin{gather*}
 u_0(0,y) = q_0,\quad u_0(1,y) = q_1,\quad u_0(x,0) = q_2,\quad u_0(x,1) = q_3,\\
 u_1 = q_4^3.
\end{gather*}
The data is chosen as follows:
\begin{align*}
 f &= \left(\begin{matrix}20\pi^2  \sin( \pi x) \sin(\pi y)\\1 \end{matrix}\right)\\
 u^d&= \left(\begin{matrix}\sin( \pi x) \sin(\pi y)*x\\x \end{matrix}\right)
\end{align*}
with $\alpha = 10$.
\subsubsection{Program description}
The control in the Dirichlet boundary values is incorporated via the
class \texttt{LocalDirichletData} which is defined in the
file \textit{localdirichletdata.h}. The class is then given to
the \texttt{OptProblemContainer} as a template argument. This is all that is
needed to use the control in the Dirichlet boundary values. 
\twick{
In the main file and the localpde program, we work still with the 
\texttt{FE\underline{}Nothing} element to assign the five controls.
}

### 2013-12-10 Examples/OPT/StatPDE/Example5/content.tex
Removed Typos in Documentation; OPT/StatPDE/Example1 documentation has been augmented; Introduction has been augmented

The requirement for the control and state mesh is that they have a common coarse grid, so the space time handler gets only one mesh (to ensure a common coarse grid), but this gets copied internally so that we have two seperate meshes for control and state. We can then separetely refine the mesh for control and state (see \texttt{RefineControlSpace} and  \texttt{RefineStateSpace}).

>>>>

The requirement for the control and state mesh is that they have a common coarse grid, so the space time handler gets only one mesh (to ensure a common coarse grid), but this gets copied internally so that we have two separate meshes for control and state. We can then separately refine the mesh for control and state (see \texttt{RefineControlSpace} and  \texttt{RefineStateSpace}).

### 2013-12-10 Examples/OPT/StatPDE/Example9/content.tex
Removed Typos in Documentation; OPT/StatPDE/Example1 documentation has been augmented; Introduction has been augmented

\begin{figure}[h]
\centering
    {\includegraphics[width=7cm]{Documentation/visit_fsi_1_opt_control_p3_0002_bearbeitet_Aug_15.png}}
    {\includegraphics[width=7cm]{Documentation/visit_fsi_1_opt_control_p3_0005_bearbeitet_Aug_15.png}}
  \caption{Configuration of the fsi cylinder-beam-drag minimization problem. 
By sucking out the fluid, 
the force on the cylinder is reduced. At left, the $x$-velocity field 
is shown. In right figure, we the corresponding adjoint solution is shown.}
  \label{ex_2_plots}
\end{figure}

@@

The rest of the progam is similar to the previous optimization problems where
we formulate the state equation in a weak form $a(q,U)(\phi)$ such that the 
final problem reads
\[
K(q,U) \rightarrow \text{min} \quad \text{s.t.} \quad a(q,U)(\phi) = 0.
\]

>>>>

\begin{figure}[h]
\centering
    {\includegraphics[width=7cm]{Documentation/visit_fsi_1_opt_control_p3_0002_bearbeitet_Aug_15.png}}
    {\includegraphics[width=7cm]{Documentation/visit_fsi_1_opt_control_p3_0005_bearbeitet_Aug_15.png}}
  \caption{Configuration of the FSI cylinder-beam-drag minimization problem. 
By sucking out the fluid, 
the force on the cylinder is reduced. At left, the $x$-velocity field 
is shown. In right figure, we the corresponding adjoint solution is shown.}
  \label{ex_2_plots}
\end{figure}

@@

The rest of the program is similar to the previous optimization problems where
we formulate the state equation in a weak form $a(q,U)(\phi)$ such that the 
final problem reads
\[
K(q,U) \rightarrow \text{min} \quad \text{s.t.} \quad a(q,U)(\phi) = 0.
\]

### 2013-12-10 Examples/PDE/InstatPDE/Example1/content.tex
Removed Typos in Documentation; OPT/StatPDE/Example1 documentation has been augmented; Introduction has been augmented

After discretization in time, the space is treated, as ususally, with 
a Galerkin finite element scheme, here based on the Taylor-Hood element 
$Q_2^c / Q_1^c$.

@@

\begin{verbatim}
void ElementEquation (..., double scale, double scale_ico)
\end{verbatim}
The two arguments 
are used to distigunish between explicit 
components and fully implicit components. 
For standard equations (such as the heat 
equation and the wave equation), there is 
no special treatment required needed.

@@

\begin{verbatim}
void ElementMatrix (..., double scale, double scale_ico)
\end{verbatim}
The directional derivatives of the state equation
are implemented in the present function. As before,
the last two parameters
\begin{verbatim}
double scale, double scale_ico
\end{verbatim} 
are used to distigungish between fully implicit 
are other behavior.

@@

In contrast to this behavior, the user 
has the possibility to write all terms of 
$\partial_t v$ explicitly. In this case, we 
use the 
\begin{verbatim}
void ElementTimeEquationExplicit (...) 
\end{verbatim}
and we write 
\begin{equation*}
(v^n - v^{n-1}, \phi)_{\Omega}.
\end{equation*}
This behavior is useful for multi-phyics problems
where other solutions variables have to be considered
around $\partial_t v$. The user should have a look 
in the second Example 
\ref{PDE_Instat_FSI} for nonstationary problems 
for an illustration
of this function.

>>>>

After discretization in time, the space is treated, as usually, with 
a Galerkin finite element scheme, here based on the Taylor-Hood element 
$Q_2^c / Q_1^c$.

@@

\begin{verbatim}
void ElementEquation (..., double scale, double scale_ico)
\end{verbatim}
The two arguments 
are used to distinguish between explicit 
components and fully implicit components. 
For standard equations (such as the heat 
equation and the wave equation), there is 
no special treatment required needed.

@@

\begin{verbatim}
void ElementMatrix (..., double scale, double scale_ico)
\end{verbatim}
The directional derivatives of the state equation
are implemented in the present function. As before,
the last two parameters
\begin{verbatim}
double scale, double scale_ico
\end{verbatim} 
are used to distinguish between fully implicit 
are other behavior.

@@

In contrast to this behavior, the user 
has the possibility to write all terms of 
$\partial_t v$ explicitly. In this case, we 
use the 
\begin{verbatim}
void ElementTimeEquationExplicit (...) 
\end{verbatim}
and we write 
\begin{equation*}
(v^n - v^{n-1}, \phi)_{\Omega}.
\end{equation*}
This behavior is useful for multi-physics problems
where other solutions variables have to be considered
around $\partial_t v$. The user should have a look 
in the second Example 
\ref{PDE_Instat_FSI} for nonstationary problems 
for an illustration
of this function.
